Our Galactic Midplane Is Probably Wrong, and Not by a Little

### 2019-01-14 paper/ms.tex
Expand intro a bit

Since the Milky Way is not truly axisymmetric, it is not clear the best way to
define the Galactocentric coordinate system --- or, perhaps, the best way to
modify the current implicit definition of the Galactocentric coordinate system
in order to account for non-axisymmetries.

In a slowly-evolving axisymmetric potential, the actions of stars' orbits are
conserved quantities \citep{2008gady.book.....B,2014RvMP...86....1S}. In other
words, if we could measure the actions of a star in such a potential, we would
measure the same actions over the course of the orbit. The non-axisymmetries
of a more realistic potential induce time evolution in the {\em measured}
actions --- the true orbital invariants (if they exist) are invariant by
definition. Through simple tests, we will see that even in a truly
axisymmetric potential, if the wrong coordinate system is assumed then there
will be additional time evolution in the measured actions.

>>>>

We define the ``global Galactic midplane'' to be the plane of $z$ symmetry in
the axisymmetric model of the Milky Way. We define the ``local Galactic
midplane'' to be the plane of $z$ symmetry inferred by an observer at a
particular location. The true global Galactic midplane is the plane of $z$
symmetry in the best-fit axisymmetric model. It is not guaranteed that the
local Galactic midplane coincide with the global Galactic midplane. Even
without a high-precision definition and measurement of the global Galactic
midplane, we can still assess how different the two should be by measuring the
azimuthal dependence of the local Galactic midplane. This procedure may be
readily tractable with current datasets. However, we turn to simulation in
this work.

It's important for us to emphasize that when we state our local Galactic
midplane may be erroneous, we do not mean that past measurements/estimates of
it are wrong. Rather, what we mean is that it may vary from the local midplane
at other points in the Galaxy. In the face of {\em Gaia} DR2 (and soon DR3/4),
this may be an important effect to consider when performing dynamical
modelling across large distances.

Since an azimuthal-dependent midplane is effectively a departure from a perfectly axisymmetric potential, we In a slowly-evolving axisymmetric potential, the actions of stars' orbits are
conserved quantities \citep{2008gady.book.....B,2014RvMP...86....1S}. In other
words, if we could measure the actions of a star in such a potential, we would
measure the same actions over the course of the orbit. The non-axisymmetries
of a more realistic potential induce time evolution in the {\em measured}
actions --- the true orbital invariants (if they exist) are invariant by
definition. Through simple tests, we will see that even in a truly
axisymmetric potential, if the wrong coordinate system is assumed then there
will be additional time evolution in the measured actions.
